Modeling the Graphotactics of Low-Resource Languages Using Sequential GANs

### 2022-05-10 tex/paper.tex
inflection training done

\documentclass{article}
\usepackage[preprint]{neurips_2022}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\bibliographystyle{plain}

@@

  \section{Method}

  \bibliography{sources}

>>>>

\documentclass{article}
\usepackage[preprint]{neurips_2022}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}
\usepackage{float}
\bibliographystyle{plain}

@@

  \section{Method}
    \subsection{Data Pipeline}
      All language data used was sourced from the SIGMORPHON 2019 repository and was initially formatted according the the Unimorph schema. The data for each language was gathered from their respective Wikipedia. Only the "train-low" sets were used for training. These datasets has only 100 examples, with each containing a lemma (base form), an inflected form, and a list of the morphological categories associated with the inflected form.\footnote{Although the "train-low" sets do occasionally have 1000 examples, those corresponding to the languages used in this experiment have only 100.}

      To produce training data for the GAN, base forms were aligned with inflected forms using a technique developed by Mans Hulden called Chinese restaurant process alignment \cite{crpalign}. Based on the alignment, a series of likely stems are generated using code from Anastasopoulos and Neubig, 2019 \cite{CMU}; as most of the languages tested use primarily suffixational inflection, the first of these "likely" stems was assumed to be the true stem and was included in the language's training set as an example of graphotactic compliance.

      Non-alphabetic symbols, such as punctuation, numerals, and spaces were removed from each stem. Although the very small training sets provided by SIGMORPHON were likely hand gathered, the data was not free of anomolous examples. While attempting to gauge the model's behavior on a familiar language, it was found that the English dataset contained extremely uncommon characters such as ``æ'', ``é'', ``ë'', ``ö'', and ``œ''; however, it was decided that was not necessary to exclude these characters, as the model should recognize their infrequency, even in a dataset of only 100 examples.

      Finally, the stems were padded to reach a length of 10 and encoded as 10 $\times$ $len(\textrm{alphabet})$ one-hot sequences.

    \subsection{Stem GAN}
      Although proven methods exist for building GANs for sequence data such as SeqGAN \cite{seqgan}, it was decided that these models would likely be too complex and therefore prone to overfitting and vanishing gradients when trained on such a small dataset. Instead, a minimalally complex GAN was created (Fig. \ref{fig:architecture}). The generator took noise input of shape 10 (representing the maximum expected stem length) $\times$ $len(\textrm{alphabet})$ and consisted of two 100 unit LSTMs with $tanh$ activations, separated by a 20\% dropout layer, and concluded with a single $len(\textrm{alphabet})$ unit dense layer with softmax activation, producing an output of equal shape to the input. The discriminator receives this output concatenated with one real example from the training set. Due to signs of mode collapse in early tests with two layers, the discriminator was given just one 100 unit LSTM. The discriminator terminates with a sigmoid activated single unit dense layer, that outputs a predicted probability that each example is real.

      \begin{figure}[H]
        \center
        \includegraphics[width=\linewidth]{architecture.pdf}
        \caption{Architecture of stem the generation model}
        \label{fig:architecture}
      \end{figure}

      The model was trained with a batch size of 1 for 1500 epochs on each low-resource language. After early tests resulted in mode mode collapse, the loss functions of both the discriminator and GAN as a whole were switched from binary cross entropy to Wasserstein loss \cite{wgan}.

    \subsection{Augmentation and Output Cleaning}
      The resultant models were each used to produce 10,000 artificial training examples. This process involved a modified version of the augmentation script from Anastasopoulos and Neubig, 2019 \cite{CMU}.

      The strings produced by the resultant models were, unfortunately, not always of the correct shape. Some had zeroes (representing null characters) in the middle, others filled the entire 10 available characters (which most real stems do not). To combat this, model output strings were "cleaned". Strings were cut off at the first zero from the left (all languages were left-to-right), if a character was repeated more than twice consecutively, it was collapsed to just two, and if the length of the stem was longer than the real stem it was replacing, it was shortened to a matching length.

    \subsection{Testing}

  \section{Results}
  \section{Discussion}

  REMEMBER TO INCLUDE LINKS TO ALL CODE CREATED AND USED
  \bibliography{sources}
